---
title: "pml_proj1"
author: "Courseraian"
date: "October 26, 2014"
output: html_document
---

First load all the libraries
#install.packages("randomForest","rpart","caret")
```{r, echo=TRUE}
library (randomForest)
library(rpart)
library(caret)
training<-read.csv("pml-training.csv", stringsAsFactor=FALSE)
```
Check the data structure

``` {r, echo=TRUE, eval=FALSE}
stdim(training)
str(training)
table(training$clsse)
```
Check those columns with data na and remove those from the data set


``` {r, echo=TRUE}
na_test <- sapply (training, function (x) {sum(is.na(x))}) 
table(na_test)  # to 
na_cols <- names(na_test[na_test==19216])
  
train_prep1 <- training[,!names(training) %in% na_cols]
```
Also Remove rown sequence and user names

```{r,echo=TRUE}
train_prep2 <- train_prep1[3:93]    
dim(train_prep2)
```
Check near Zero variable to identify those without changes. they may not good predictors. 
```{r, echo=TRUE}
nearzero <- nearZeroVar(train_prep2, saveMetrics=T)
names(nearzero)
sum(nearzero$nzv)

```
Remove near zero cols
```{r}
# need to do the same for testing data?
train_nz <- train_prep2[, nearzero$nzv]
train_rest <- train_prep2[, !nearzero$nzv]
```
Partition training set to two diff groups: train_trn is used to build a model, and train_val is used to validate it. 

```{r}
train_rest_indx <-createDataPartition(y=train_rest$classe, p =0.7, list = FALSE) 
train_trn <- train_rest[train_rest_indx,]
train_val <- train_rest[-train_rest_indx,]
```

Build models
Warning: This part was messed up at last minutes when I tried different models. I run out of time. 
```{r, echo=TRUE, eval=FALSE}
set.seed(4321)


rfMod <- train(train_trn[,-1], train_trn$classe, preProcess = c("center", "scale"),trControl=trainControl(method="boot"))

summary(rfMod )
print(rfMod)
importance(rfMod)
imp <- importance(rfMod)
print(importance(rfMod))
plot(rfMod)
plot(importance(rfMod),lty=2, pch=16)
#imp_sort <- imp[order(imp[,1],decreasing=TRUE)]
lines(importance(rfMod))
pred_val <- predict (rfMod, newdata=train_val_comp, type="response")
train_conf <- confusionMatrix(train_val_comp$classe, pred_val)
varImp(rfMod) # variale impor




rfMod <- randomForest(train_trn[,-57], train_trn$classe )

summary(rfMod )
print(rfMod)
importance(rfMod)
imp <- importance(rfMod)
print(importance(rfMod))
plot(rfMod)
plot(importance(rfMod),lty=2, pch=16)
#imp_sort <- imp[order(imp[,1],decreasing=TRUE)]
lines(importance(rfMod))
pred_val <- predict (rfMod, newdata=train_val_comp, type="response")
train_conf <- confusionMatrix(train_val_comp$classe, pred_val)
varImp(rfMod) # variale importance. same with importance. 

```
Now bring in test data and be sure to preprocess the test data the same way you preocess the training data
```{r, echo=TRUE, eval=FALSE}
testing <-read.csv("pml-testing.csv")
test_prep2 <- testing[,!names(testing) %in% na_cols]

test_nearzero <- nearZeroVar(testing, saveMetrics=T)
test_nz <- testing[, nearzero$nzv]
test_rest <- testing[, !nearzero$nzv]
test_prep1 <- test_rest[,6:length(test_rest)]
test_comp <- test_prep1[complete.cases(test_prep1),]

pred_test <- predict (rfMod, newdata=test_comp, type="response")

test_conf <- confusionMatrix(test_comp$classe, pred_test)

varImp(rfMod)
```


```{r,echo=FALSE, eval=FALSE}
# remove high correlation







tt_prep <- preProcess(train_trn[,-91], method=c("center","scale"), na.remove=TRUE)

#tt_prcomp <- prcomp(train_trn[,-95])


####
# train_trn_comp <- train_trn[complete.cases(train_trn),] 
# train_val_comp <- train_val[complete.cases(train_val),]




#try rpart
rpartMod <- train(classe ~., data = train_trn)
summary(rpartMod )
plot(rpartMod$finalModel)
rp_pred_val <- predict (rpartMod, newdata=train_val)
summary(rp_pred_val)
rp_train_conf <- confusionMatrix(train_val$classe, rp_pred_val)

# try random forest
rfMod <- randomForest(classe ~., data = na.omit(train_trn))
summary(rfMod )
print(rfMod)
importance(rfMod)
imp <- importance(rfMod)
print(importance(rfMod))
plot(rfMod)
plot(importance(rfMod),lty=2, pch=16)
#imp_sort <- imp[order(imp[,1],decreasing=TRUE)]
lines(importance(rfMod))
pred_val <- predict (rfMod, newdata=train_val_comp, type="response")
train_conf <- confusionMatrix(train_val_comp$classe, pred_val)
varImp(rfMod) # variale importance. same with importance. 



# remove high correlation



# try lda

ldaMod <- train (classe~., data = train_trn, method='lda')

# try glm

ldaMod <- train (classe~., data = train_trn,method='glm')


preProcess





